The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2018-08-30 rnw/Introduction.Rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

In many countries, forest inventories have become an indispensable tool for evaluating the current state of forests as well as for tracking their development over time. They provide accurate quantitative information that can be used to define management actions and to adapt forest management strategies according to guidelines on national and international levels. As conducting a full census of all trees within any sizable forest area is clearly infeasible due to time and cost restrictions, forest inventories usually gather their information by means of statistical sampling methods. Typically this means that discrete sample locations (sample plots) are randomly chosen in the forest, making up the framework of a terrestrial inventory.  This terrestrial sample data is then used to make estimates for the entire forested area and provide a measure of precision for those estimates in the form of confidence intervals. There is a broad range of literature describing the concepts and methods regarding the choice of different estimators under various sample designs \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.

Terrestrial inventories have the benefit of being very flexible in the sense that they can be used to produce high quality estimates for a wide-variety of different forest attributes. However, they have the downside of being very expensive.  Improving the precision of the estimates by increasing the number of sample plots essentially means that travel costs will rise as trained inventorists are sent to more and more plot locations. This is why the number of terrestrial samples is often limited. Although national inventories usually possess a sufficiently large terrestrial sample size to provide high estimation accuracies for larger areas, this is often not the case for smaller areas, such as forest management units.  As a result, there has been an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}. A method which has become particularly attractive is called multi-phase sampling. The core concept is to enlarge the sample size in order to gain higher estimation precision without enlarging the terrestrial sample size. This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered. These predictions are produced by regression models that use explanatory variables derived from auxiliary
data, commonly in the form of spatially exhaustive remote sensing data in the inventory area. Regression estimators using this concept can consider either one additional sample of plot locations (two-phase or double-sampling) or two additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}. Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par

% section: briefly give a review of existing R-packages for forest inventories (not much around)
% While commercial enterprises have already been offering multi-phase inventory analyses, freely available software for application of multi-phase estimators has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available. The motivation behind the \proglang{R} package \pkg{forestinventory} is to offer scientists and forest practitioners a comprehensive set of multi-phase estimation procedures that provide the advantages of a) full transparency of the implemented estimation methods, b) free access also for (small) forest enterprises who cannot afford commercial services, and c) implementation in an open source software that provides high flexibility to be incorporated in operational forest information systems.\par
% 
% The \pkg{forestinventory} package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management. The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.

Multi-stage and multi-phase estimation has already been implemented in commercial as well as open-source software, such as the survey sampling procedures in \proglang{SAS} \citep{sas} and the \pkg{survey} package in \proglang{R} \citep{survey}. However, both are targeted towards list-sampling as it is applied in official statistics. Available software providing multi-phase sampling methods better suited for forest inventories has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the \pkg{maSAE} package by \citet{cullmann2016}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

@@

% Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods to forestry has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

>>>>

In many countries, forest inventories have become an indispensable tool for evaluating the current state of forests as well as for tracking their development over time. They provide accurate quantitative information that can be used to define management actions, to adapt forest management strategies according to guidelines on national and international levels and to support in monitoring carbon change in the context of climate reporting and mitigation. As conducting a full census of all trees within any sizable forest area is clearly infeasible due to time and cost restrictions, forest inventories usually gather their information by means of statistical sampling methods. Typically this means that discrete sample locations (sample plots) are randomly chosen in the forest, making up the framework of a terrestrial inventory.  This terrestrial sample data is then used to make estimates for the entire forested area and provide a measure of precision for those estimates in the form of confidence intervals. There is a broad range of literature describing the concepts and methods regarding the choice of different estimators under various sample designs \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.

Terrestrial inventories have the benefit of being very flexible in the sense that they can be used to produce high quality estimates for a wide-variety of different forest attributes. However, they have the downside of being very expensive.  Improving the precision of the estimates by increasing the number of sample plots essentially means that travel costs will rise as trained inventorists are sent to more and more plot locations. This is why the number of terrestrial samples is often limited. Although national inventories usually possess a sufficiently large terrestrial sample size to provide high estimation accuracies for larger areas, this is often not the case for smaller areas, such as forest management units.  As a result, there has been an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}. A method which has become particularly attractive is called multiphase sampling. The core concept is to enlarge the sample size in order to gain higher estimation precision without enlarging the terrestrial sample size. This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered. These predictions are produced by regression models that use explanatory variables derived from auxiliary
data, commonly in the form of spatially exhaustive remote sensing data in the inventory area. Regression estimators using this concept can consider either one additional sample of plot locations (two-phase or double-sampling) or two additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}. Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par

% section: briefly give a review of existing R-packages for forest inventories (not much around)
% While commercial enterprises have already been offering multiphase inventory analyses, freely available software for application of multiphase estimators has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available. The motivation behind the \proglang{R} package \pkg{forestinventory} is to offer scientists and forest practitioners a comprehensive set of multiphase estimation procedures that provide the advantages of a) full transparency of the implemented estimation methods, b) free access also for (small) forest enterprises who cannot afford commercial services, and c) implementation in an open source software that provides high flexibility to be incorporated in operational forest information systems.\par
% 
% The \pkg{forestinventory} package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management. The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.

Multistage and multiphase estimation has already been implemented in commercial as well as open-source software, such as the survey sampling procedures in \proglang{SAS} \citep{sas} and the \pkg{survey} package in \proglang{R} \citep{survey}. However, both are targeted towards list-sampling as it is applied in official statistics. Available software providing multiphase sampling methods better suited for forest inventories has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the \pkg{maSAE} package by \citet{cullmann2016}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

@@

% Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multiphase sampling methods to forestry has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

### 2018-08-30 rnw/Main structure and modules of the Package.Rnw
update 06_08_2018
update 06_08_2018 - first hands on revision
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by coordinates. The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g., timber volumes) to local plot densities (e.g., the timber density in m$^3$/ha). The estimators implemented in \pkg{forestinventory} use the so called infinite population approach in order to bridge this inventory process to the mathematics behind the estimators. This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Fig. \ref{fig:inf_pop_apr}. The population total of the target variable (e.g., the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Fig. \ref{fig:inf_pop_apr} as the volume under the density surface. From this perspective, the practical challenge is that the form of this function is unknown. Theoretically, we could get the total timber volume by observing the function value, i.e., the local density $Y(x)$, at each possible point $x$ over the forest area and taking their sum. However, this is impossible because there is an infinite number of points in the forest area. Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities $Y(x)$ to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.  The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.  All estimators included in \pkg{forestinventory} rest upon this theoretical perspective. The key point in the infinite population approach is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area. This has some theoretical advantages over the finite population approach, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.  This is mainly due to the impossibility of a perfect tessellation over an amorphous forest area by any choice of plot shape. Hence, the population in the finite approach is, strictly speaking, not well defined with respect to the forest area. The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.

@@

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{1\hsize}{!}{\includegraphics{fig/phases_graphic(new).eps}}%
		\caption{} \label{fig:concmphase_and_sae_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
	 \resizebox{0.58\hsize}{!}{\includegraphics{fig/sae_graphic.PNG}}%
		\caption{} \label{fig:concmphase_and_sae_b}
	\end{subfigure}
\caption{(a) Concept of multi-phase sampling. The square represents the forest area for which an inventory is being conducted. The points denote the sample locations $x$. Filled points indicate available information. (b) Illustration of the small area estimation problem.}
\label{fig:concmphase_and_sae}
\end{figure}

@@

Small area estimation does not necessarily refer to small spatial areas but rather to areas that contain little or no terrestrial sample. To formulate this mathematically, we want to make an estimate for a subregion $G$ of the entire inventory area $F$ (Fig. \ref{fig:concmphase_and_sae_b}). As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multi-phase estimation can be efficient. However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals. The idea is then to borrow strength from the entire terrestrial sample $s_2$ of $F$ to fit the model, and to apply this model to the small area. The potential bias of applying that model in $G$ is then corrected for by using the empirical model residuals derived from that small area. If there are no terrestrial plots in $G$ (i.e., $n_{2,G}=0$), one cannot correct for a potential model bias in $G$ and has to accept a potential bias in the estimator. These are called synthetic estimates and despite their potential bias, it is usually still possible to calculate their design-based variance.

@@

In the \pkg{forestinventory} package, estimators for two-phase and three-phase sampling are applied with the \code{twophase()} and \code{threephase()} functions. From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Fig. \ref{fig:struct_package}). Choosing an estimator follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications. The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (global estimators) or only for subregions of the entire inventory area (small area estimators). In the second case, the package offers three small area estimators that will in detail be described in the following sections. The estimators are available under exhaustive and non-exhaustive use of the auxiliary data. Additionally, the package can also calculate one-phase estimates solely based on terrestrial samples. All estimators are also available for cluster sampling, in which case a sample unit consists of multiple, spatially agglomerated samples. The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}. While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package. We will also restrict discussion to simple sampling, while the formulas for cluster sampling are available in the technical reports \citep{mandallaz2016, mandallaz2013techa, mandallaz2013techb}. A special case under cluster sampling is described in Section \ref{sec:speccas_and_scen}.

\begin{figure}[htb]
\centering
\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_5.eps}}
\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}.}
\label{fig:struct_package}
\end{figure}

>>>>

Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by coordinates. The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g., timber volumes) to local plot densities (e.g., the timber density in m$^3$/ha) including potential boundary adjustments (see \cite{mandallaz2008} for details). The estimators implemented in \pkg{forestinventory} use the so-called infinite population approach in order to bridge this inventory process to the mathematics behind the estimators. This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Fig. \ref{fig:inf_pop_apr}. The population total of the target variable (e.g., the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Fig. \ref{fig:inf_pop_apr} as the volume under the density surface. From this perspective, the practical challenge is that the form of this function is unknown. Theoretically, we could get the total timber volume by observing the function value, i.e., the local density $Y(x)$, at each possible point $x$ over the forest area and taking their sum. However, this is impossible because there is an infinite number of points in the forest area. Our strategy is thus to take a sample of $n_2$ points, $s_2$, from an infinite population of possible points and use their associated local densities $Y(x)$ to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.  The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.  All estimators included in \pkg{forestinventory} rest upon this theoretical perspective. The key point in the infinite population approach is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area. Sampling the actual plot areas follows the so-called finite population approach, where the sample units are usually assumed to be either circular or rectangular plots.  Although choosing between infinite and finite population schemas does not likely make much practical difference, the former does has some theoretical advantages.  This is mainly due to the impossibility of a perfect tessellation over an amorphous forest area by any choice of plot shape. Hence, the population in the finite approach is, strictly speaking, not well defined with respect to the forest area.  Also, cluster sampling is much more easily generalized in the infinite population approach. The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.

@@

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{1\hsize}{!}{\includegraphics{fig/phases_graphic(new).eps}}%
		\caption{} \label{fig:concmphase_and_sae_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
	 \resizebox{0.58\hsize}{!}{\includegraphics{fig/sae_graphic.PNG}}%
		\caption{} \label{fig:concmphase_and_sae_b}
	\end{subfigure}
\caption{(a) Concept of multiphase sampling. The square represents the forest area for which an inventory is being conducted. The points denote the sample locations $x$. Filled points indicate available information. (b) Illustration of the small area estimation problem.}
\label{fig:concmphase_and_sae}
\end{figure}

@@

Small area estimation does not necessarily refer to small spatial areas but rather to areas that contain little or no terrestrial sample. To formulate this mathematically, we want to make an estimate for a subregion $G$ of the entire inventory area $F$ (Fig. \ref{fig:concmphase_and_sae_b}). As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multiphase estimation can be efficient. However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals. The idea is then to borrow strength from the entire terrestrial sample $s_2$ of $F$ to fit the model, and to apply this model to the small area. The potential bias of applying that model in $G$ is then corrected for by using the empirical model residuals derived from that small area. If there are no terrestrial plots in $G$ (i.e., $n_{2,G}=0$), one cannot correct for a potential model bias in $G$ and has to accept a potential bias in the estimator. These are called synthetic estimates and despite their potential bias, it is usually still possible to calculate their design-based variance.

@@

In the \pkg{forestinventory} package, estimators for two-phase and three-phase sampling are applied with the \code{twophase()} and \code{threephase()} functions. From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Fig. \ref{fig:struct_package}). Choosing an estimator follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications. The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (global estimators) or only for subregions of the entire inventory area (small area estimators). In the second case, the package offers three small area estimators that will in detail be described in the following sections. The estimators are available under exhaustive and non-exhaustive use of the auxiliary data. Additionally, the package can also calculate one-phase estimates solely based on terrestrial samples. All estimators are also available for cluster sampling, in which case a sample unit consists of multiple, spatially agglomerated samples. The following sections describe the mathematical details and the application of the multiphase estimators implemented in the \proglang{R} package \pkg{forestinventory}. While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package. We will also restrict discussion to simple sampling, while the formulas for cluster sampling are available in the technical reports \citep{mandallaz2016, mandallaz2013techa, mandallaz2013techb}. A special case under cluster sampling is described in Section \ref{sec:speccas_and_scen}.

\begin{figure}[htb]
\centering
\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_5.eps}}
\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}. All estimators are also available for cluster sampling.}
\label{fig:struct_package}
\end{figure}

### 2018-08-30 rnw/Onephase_Estimators and their Application.Rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

>>>>

In the package \pkg{forestinventory}, calculating estimates exclusively based on terrestrial observations, i.e. one-phase estimations, can be applied with the \code{onephase()} function. To demonstrate the function, we will use the \code{grisons} data frame that comes with installing the package from the CRAN repository. The dataset contains data from a two-phase forest inventory under simple sampling design conducted in 2007 that was used in \citet{mandallaz2013b} and comprises observations from 306 sample locations ($s_1$ sample). A systematic subsample of 67 out of the 306 observations ($s_2$ sample) provides the local density $Y(x)$, in this case the timber volume per hectare (\code{tvol}) derived from the terrestrial inventory campaign. We can load the data frame by executing:

\begin{small}
<<>>=
data("grisons", package = "forestinventory")
@
\end{small}

The inventory data for the \code{onephase} estimation must be provided as a \code{data.frame} that contains a column with the local density values and additionally a column specifying to which sample ($s_0$, $s_1$, $s_2$) the respective observation belongs. In the simple one-phase case, all observations in the data frame will of course only belong to one, i.e. the terrestrial phase. However, the specification is mandatory since in most applications, the input dataset will contain observations from multiple phases, and the indication of the terrestrial sample is thus required to identify the terrestrial sample.

We apply the \code{onephase()} function by specifying the following arguments:

\begin{itemize}
  \itemsep0em
  \item the name of the column containing the local density $Y$, which is defined as a formula expression that must be of the form $Y~1$
  \item the \code{data.frame} containing the inventory data
  \item the \code{list}-object \code{phase_id} taking the column name (\code{phase.col}) specifying the sample association of an observation ($s_1$ or $s_2$) and the indicator for the $s_2$ phase (\code{terrgrid.id})
\end{itemize}

\begin{small}
<<>>=
op <- onephase(formula = tvol~1 ,data = grisons,
              phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))y
@
\end{small}

### 2018-08-30 rnw/Special Cases and Scenarios.Rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

A special case of multi-phase regression estimation is post-stratification, which can further be divided into the cases of multi-phase sampling for stratification and multi-phase sampling for regression within strata. Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.

@@

Two-phase sampling for stratification is applied if the model only contains categorical variables, in this case the factor variable \code{stage}. Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multi-phase regression estimators, is equivalent to post-stratification. For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).

@@

Two-phase sampling for regression within strata implies the combination of continuous and categorical variables within the model (i.e., we have an ANCOVA model). If an interaction term is not present between categorical and continuous variables, the regression lines within the strata will have the same slope but different intercepts. If an interaction term is present, both the intercept and the slope are allowed to vary within the strata. Note that one can actually use the entire range of OLS regression techniques in the multi-phase estimators, including higher order terms and transformations of the explanatory variables, which makes them very flexible.

@@

As explained in Section \ref{sec:str_and_mod}, a basic prerequisite for the application of multi-phase estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are nested in each other. The correct nesting thereby concerns the spatial arrangement of the sample phases (Fig. \ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location. For the latter, \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first-aid adjustments to the inventory data set to prevent the calculations from failing. We will demonstrate possible nesting violations by applying the global three-phase estimator to the \code{grisons} and \code{zberg} data sets.

@@

Based on the nesting rule, $s_2 \in s_1 \in s_0$, each $s_2$ and $s_1$ sample location must have all explanatory variables available that are used in the full (and thus reduced) regression model. If e.g., an $s_2$ and/or $s_1$ point misses a variable which is used in the full and reduced model (in this case \code{mean}), the function will delete this sample point from the dataset and produce the following messages:

>>>>

A special case of multiphase regression estimation is post-stratification, which can further be divided into the cases of multiphase sampling for stratification and multiphase sampling for regression within strata. Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.

@@

Two-phase sampling for stratification is applied if the model only contains categorical variables, in this case the factor variable \code{stage}. Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multiphase regression estimators, is equivalent to post-stratification. For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).

@@

Two-phase sampling for regression within strata implies the combination of continuous and categorical variables within the model (i.e., we have an ANCOVA model). If an interaction term is not present between categorical and continuous variables, the regression lines within the strata will have the same slope but different intercepts. If an interaction term is present, both the intercept and the slope are allowed to vary within the strata. Note that one can actually use the entire range of OLS regression techniques in the multiphase estimators, including higher order terms and transformations of the explanatory variables, which makes them very flexible.

@@

As explained in Section \ref{sec:str_and_mod}, a basic prerequisite for the application of multiphase estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are nested in each other. The correct nesting thereby concerns the spatial arrangement of the sample phases (Fig. \ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location. For the latter, \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first-aid adjustments to the inventory data set to prevent the calculations from failing. We will demonstrate possible nesting violations by applying the global three-phase estimator to the \code{grisons} and \code{zberg} data sets.

@@

Based on the nesting rule, $s_2 \subset s_1 \subset s_0$, each $s_2$ and $s_1$ sample location must have all explanatory variables available that are used in the full (and thus reduced) regression model. If e.g., an $s_2$ and/or $s_1$ point misses a variable which is used in the full and reduced model (in this case \code{mean}), the function will delete this sample point from the dataset and produce the following messages:
